Black-hole ringdown as a probe of higher-curvature gravity theories

### 2022-04-08 papers/paper_alt_theor_bounds/paper_alt_theor_bounds.tex
Section V.A

We start by showing our results for EdGB gravity.
%
In Fig.~\ref{fig:edgb_exec_sum} we show the marginalized posteriors of the
coupling constant $\ell_{\rm GB}$, for GW150914 (top panel) and GW200129
(middle panel), with and without the spin corrections to the $(2,2,0)$ QNM.
%
The bottom panel panel shows the joint posterior combining both events.
%
We see that for both events events, the $N_{\rm max} = 0$ posteriors are
characterized by a peak away from zero.
%
This does not mean that we are inferring a deviation from GR from data.
%
Recall that the deviations from GR in the ParSpec framework are controlled
by $\gamma$, which here reads,
%
\begin{equation}
    \gamma_{\rm GB} = \left(\frac{c^2 \ell_{\rm GB}}{ G M_{f}}\right)^4\,.
    \label{eq:def_gamma_gb}
\end{equation}
%
As shown in Fig.~\ref{fig:gamma_gb}, $\gamma_{\rm GB}$ does indeed have a posterior
distribution with largest support at zero, indicating consistency between data
and GR.
%
We also observe that the inclusion of the spin corrections (i.e., curves with $N_{\rm max} = 1$)
displaces the posteriors distributions towards smaller values of $\ell_{\rm GB}$.

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/edgb_posteriors_combined.pdf}
\caption{(Color online).~Posterior distribution function in EdGB gravity
for the coupling constant $\ell_{\rm GB}$ for GW150914 (top panel), GW200129 (middle panel)
and combined events (bottom panel).
%
In all panels, different line colors correspond to the inclusion ($N_{\rm max} = 1$) or not ($N_{\rm max} = 0$)
of the linear-in-spin QNM correction.
%
The joint posteriors are shown for illustrative purposes only. As we explain in Fig.~\ref{fig:edgb_cdf} and
in the main text, our analysis of these events fails to satisfy the EFT bound~\eqref{eq:eft_bound}.
}
\label{fig:edgb_exec_sum}
\end{figure}

@@

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/edgb_cdf_varying_threshold.pdf}
\caption{(Color online).~The CDF evaluated at the
cut-off $\Lambda_{\rm EFT}(\varepsilon,\gm)$ EdGB gravity as a function of the parameter
$\varepsilon$ for both (dashed curves) $\gm = m_2$, the secondary's source
mass, and (solid curves) $\gm = M_{f}$, the remnant's source mass, without
(black curves) and with (purple curves) linear in spin QNM corrections. The
horizontal lines lay at 0.68 and 0.90. We see that in no situation the curves pass
through the 0.68 or 0.90 lines even at $\varepsilon = 1$.
%
This means that no bound on $\ell_{\rm GB}$ can be placed with the events we
analyzed.
}
\label{fig:edgb_cdf}
\end{figure}

@@

As we emphasized in Sec.~\ref{sec:remarks}, we must first check whether the
``EFT''~\eqref{eq:eft_bound} and ``ParSpec''~\eqref{eq:parspec_bound} bounds
are satisfied, before drawing any conclusions on the allowed values for
$\ell_{\rm GB}$ from our parameter estimations
%
We check the validity of the EFT bound in Fig.~\ref{fig:edgb_cdf}. In the top (bottom) panel we show
the CDF of the $\ell_{\rm GB}$ posteriors for GW150914 (GW200129), obtained
by evaluating the integral~\eqref{eq:def_cdf} with $\ell_{\rm max} = \Lambda_{\rm EFT}(\varepsilon, \mathfrak{m})$,
with the mass scale set by the secondary's mass (i.e., $\mathfrak{m} = m_2$, dashed lines) or
the remnant's mass ($\mathfrak{m} = M_f$, solid lines), while varying $\varepsilon$ between 0 and 1.
%
For GW150914, we see that for the $N_{\rm max} = 0$ curves, that the CDF never goes
past $0.2$, regardless of the mass scale $\mathfrak{m}$ used and even at $\varepsilon =
1$, at which the EFT description of the theory would not be valid anyway.
%
This shows that that the ``EFT bound'' given by Eq.~\eqref{eq:eft_bound} is
never met at a significant credibility and thus that we cannot place a bound on
$\ell_{\rm GB}$.
%
The situation is similar for GW200129 with $N_{\rm max} = 0$ and does not
change for either event when we add spin corrections to the EdGB QNM.
%
Taken together we are led to \emph{conclude that we cannot constrain EdGB gravity with our
present model}.
%
We summarize our findings in Table~\ref{tab:summary_edgb}.

>>>>

We start \sout{by showing our results for} \ag{with} EdGB gravity.
%
In Fig.~\ref{fig:edgb_exec_sum} we show the marginalized PDFs of the
coupling constant $\ell_{\rm GB}$, for GW150914 (top panel) and GW200129
(middle panel), with and without the spin corrections to the $(2,2,0)$ QNM.
%
The bottom panel shows the joint posterior from combining both events.
%
We see, for both events events, the $N_{\rm max} = 0$ posteriors are
characterized by a peak away from zero.
%
This does not mean that we are inferring a deviation from GR.
%
Recall that the deviations from GR in the ParSpec framework are controlled
by $\gamma$, which here reads,
%
\begin{equation}
    \gamma_{\rm GB} = \left(\frac{c^2 \ell_{\rm GB}}{ G M_{f}}\right)^4\,.
    \label{eq:def_gamma_gb}
\end{equation}
%
As shown in Fig.~\ref{fig:gamma_gb}, $\gamma_{\rm GB}$ does indeed have a posterior
distribution with largest support at zero, indicating consistency between \sout{data} 
\ag{the underlying signal} \agcomm{whenever possible, I replaced ``data" when we 
meant signal, simply because earlier we defined data to also include noise, and by a
GR prediction we conventionally refer to the signal. }
and GR.
%
We also observe that the inclusion of the spin corrections (i.e., curves with $N_{\rm max} = 1$)
displaces the posteriors distributions towards smaller values of $\ell_{\rm GB}$ \ag{(Fig.~\ref{fig:edgb_exec_sum}).
and larger values of $\gamma_{\rm GB}$ (Fig.~\ref{fig:gamma_gb}).}

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/edgb_posteriors_combined.pdf}
\caption{(Color online).~Posterior distribution function on coupling constant $\ell_{\rm GB}$ in EdGB gravity
for GW150914 (top panel), GW200129 (middle panel)
and from combining results (bottom panel).
%
In all panels, different line colors correspond to the inclusion ($N_{\rm max} = 1$) or not ($N_{\rm max} = 0$)
of the linear-in-spin QNM correction.
%
The joint posteriors are shown for illustrative purposes only. As we explain in Fig.~\ref{fig:edgb_cdf} and
in the main text, our analysis of these events fails to satisfy the EFT bound~\eqref{eq:eft_bound}.
}
\label{fig:edgb_exec_sum}
\end{figure}

@@

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/edgb_cdf_varying_threshold.pdf}
\caption{(Color online).~The CDF evaluated at the
cut-off $\Lambda_{\rm EFT}(\varepsilon,\gm)$ for EdGB gravity as a function of the parameter
$\varepsilon$ for both (dashed curves) $\gm = m_2$, the secondary's source
mass, and (solid curves) $\gm = M_{f}$, the remnant's source mass, without
(black curves) and with (purple curves) linear in spin QNM corrections. The
horizontal lines \sout{lay at 0.68 and 0.90}\ag{mark the 68\% and 90\% credible levels}. 
We see that in no situation the curves pass
through the 0.68 or 0.90 lines even at $\varepsilon = 1$.
\agcomm{Am I missing something something? The Nmax=1 lines for GW150914 does 
cross the credible levels, as well as the m2 Nmax=1 line for GW200129.}
%
This means that no bound on $\ell_{\rm GB}$ can be placed with the events we
analyzed.
}
\label{fig:edgb_cdf}
\end{figure}

@@

As we emphasized in Sec.~\ref{sec:remarks}, we must first check whether the
``EFT''~\eqref{eq:eft_bound} and ``ParSpec''~\eqref{eq:parspec_bound} bounds
are satisfied, before drawing any conclusions on the allowed values for
$\ell_{\rm GB}$ from our parameter estimations.
%
We check the validity of the EFT bound in Fig.~\ref{fig:edgb_cdf}. In the top (bottom) panel we show
the CDF of the $\ell_{\rm GB}$ posteriors for GW150914 (GW200129), obtained
by evaluating the integral~\eqref{eq:def_cdf} with $\ell_{\rm max} = \Lambda_{\rm EFT}(\varepsilon, \mathfrak{m})$,
with the mass scale set by the secondary's mass (i.e., $\mathfrak{m} = m_2$, dashed lines) or
the remnant's mass ($\mathfrak{m} = M_f$, solid lines), while varying $\varepsilon$ between 0 and 1.
%
For GW150914, we see that for the $N_{\rm max} = 0$ curves, that the CDF never goes
past $0.2$, regardless of the mass scale $\mathfrak{m}$ used and even at $\varepsilon =
1$, at which the EFT description of the theory would not be valid anyway.
%
This shows that that the ``EFT bound'' given by Eq.~\eqref{eq:eft_bound} is
never met at a significant credibility and thus that we cannot place a bound on
$\ell_{\rm GB}$.
%
The situation is similar for GW200129 with $N_{\rm max} = 0$ and does not
change for either event when we add spin corrections to the EdGB QNM.
%
Taken together we are led to \emph{conclude that we cannot constrain EdGB gravity with our
present model}.
%
We summarize our findings in Table~\ref{tab:summary_edgb}.
